---
title: "GGPlot"
format: 
  html:
    toc: true
editor: source
---

```{r, warning=F, message=F}
library(ggplot2)
library(dplyr)
```

## Description
I downloaded corn yield data from [USDA's quick stats lite site](https://www.nass.usda.gov/Quick_Stats/Lite/index.php#91CC0BA8-1B17-3774-91BB-3549B470CD60) so that I could compare corn yields over time from a couple of states of interest: Montana (where I will be moving to) and Colorado (because of CSU).

To replicate the data one can visit the site linked above and apply the filters described below:
* Sector = CROPS
* Group = FIELD CROPS
* Commodity = CORN
* View = Acreage, Yield, and Production - Irrigated / Non-Irrigated
* Year = 1950-2022
* Geographic Level = State

**Note: ** I exported data for all states, but had to do it in two exports. The tool provided an empty CSV when I attempted to select data for all states at once.

## Load Data
```{r}
# Read data from a csv file
# data <- read.csv('corn_production.csv')
data1 <- read.csv('corn_production1.csv')
data2 <- read.csv('corn_production2.csv')
data <- bind_rows(data1, data2)

# Change the names to lower case so that they're easier for me to work with
names(data) <- tolower(names(data))

# Display the first few rows
head(data)
```
## Data preparation

Here I will rename, select a subset of columns, and filter the data to records that have data for production

```{r}
# Rename, select a subset of columns, and filter
data <- data %>%
  rename(
    production=production.in.bu, 
    harvested_area=area.harvested.in.acres,
    yield=yield.in.bu...acre
    ) %>% 
  filter(production > 0) %>% 
  select(year, location, prodn.practice, harvested_area, production, yield)

# Display the first few rows
head(data)
```

## Plotting

Here I plot the data of the states that the data was queried for

```{r}
ggplot(data, aes(x=year, y=yield, color=location)) +
  facet_wrap(~prodn.practice, ncol=2) +
  geom_line()
```

