---
title: "GGPlot Assignment"
format: html
editor: source
---

```{r, warning=F, message=F}
library(dplyr)
library(ggplot2)
library(gghighlight)
```

## Description

I downloaded corn yield data from [USDA's quick stats lite site](https://www.nass.usda.gov/Quick_Stats/Lite/index.php#91CC0BA8-1B17-3774-91BB-3549B470CD60) so that I could compare corn yields for irrigated land over time from a couple of states of interest: Montana (where I will be moving to) and Colorado (because of CSU).

To replicate the data one can visit the site linked above and apply the filters described below:

-   Sector = CROPS
-   Group = FIELD CROPS
-   Commodity = CORN
-   View = Acreage, Yield, and Production - Irrigated / Non-Irrigated
-   Year = 1950-2022
-   Geographic Level = State

**Note:** I exported data for all states, but had to do it in two exports. The tool provided an empty CSV when I attempted to select data for all states at once. This data is also available in this GitHub repo.

## Load Data

```{r}
# Read data from a csv file
# data <- read.csv('corn_production.csv')
data1 <- read.csv('corn_production1.csv')
data2 <- read.csv('corn_production2.csv')
data <- bind_rows(data1, data2)

# Change the names to lower case so that they're easier for me to work with
names(data) <- tolower(names(data))
```

## Data preparation

Here I will rename, select a subset of columns, and filter the data to records that have data for production as well as those where the corn was grown on irrigated land.

```{r}
# Rename, select a subset of columns, and filter
data <- data %>%
  rename(
    production=production.in.bu, 
    harvested_area=area.harvested.in.acres,
    yield=yield.in.bu...acre
    ) %>% 
  filter(
    production > 0,
    prodn.practice == 'IRRIGATED'
    ) %>% 
  select(year, location, prodn.practice, harvested_area, production, yield)
```

## Plotting

```{r, warning=FALSE, message=FALSE}
ggplot(data, aes(x=year, y=yield)) +
  theme_minimal() +
  geom_line(aes(group=location), color='red') +
  gghighlight(location %in% c('MONTANA', 'COLORADO')) +
  labs(
    title='USDA Corn Yield Data',
    subtitle='Comparison of Montana and Colorado corn yields',
    x='Year', 
    y='Yield (bu/acre)'
    )
```
